### 2019-06-03 Intro.Rmd
discussion

<!--INTRODUCTION: 1st paragraph: general statement that the field knows to be true and important-->
Time of lineage divergence <!--alt lineages: named biological entities, taxon ages?-->
constitutes a fundamental piece of information for evolutionary
understanding in many areas of research, from developmental to conservation biology [@Felsenstein1985a; @Webb2000]. This information can be obtained from fossils and from molecular dating studies. The developments in DNA sequencing and phylogenetic dating inferences have encouraged/spurred the application of molecular dating methods to a very large diversity of organisms, greatly increasing the quantity 
<!-- and improving the quality  (do we know this? we would expect thatm but there is no analysis on that really...)-->
of data on taxon ages across the tree of life.
Along with phylogenetic relationships, it allows testing alternative
evolutionary hypothesis in many areas of research, from developmental to conservation biology [@Felsenstein1985a; @Webb2000].
It has been shown that only a bit of information on time of divergence can improve inferences of phylogenetic distance @Webb2008 bladj
<!--comparative studies, in general, for which knowing
the time context for all life is crucial.-->
Together with geographical distribution patterns, it is a crucial component of historical biogeography studies [@posadas2006historical].
Coupled to species number, it allows studying the tempo and mode of speciation and extinction,
<!-- crucial/vital/critical/essential -->
central for understanding how biodiversity
patterns are shaped across space, time, and clades [@Morlon2014].
<!-- Efforts to have a whole tree of life have been great and here are some examples. -->
<!--INTRODUCTION: 2nd paragraph: important advances in the field-->
In the past two decades, the possibility to obtain good quality DNA sequences
<!-- from a wide variety of organisms became a reality, which,  -->
coupled to methodological developments in phylogenetic and dating inference, allowed
the application of molecular dating methods to a very large diversity
of organisms, greatly increasing the quantity and quality of<!--organisms with--> data on taxon ages across the tree of life.
<!-- figure showing growth on studies with age data from references table from ttol-->
To date, there is a large amount of
<!-- both fossil and molecular-based  -->
data on taxon ages and phylogenetic relationships in public repositories such as Dryad, TreeBASE
and Open Tree of Life (OToL).
<!-- pbdb ??-->
<!--OToL alone hosts now more than 200 chronograms. Cache updated jan 15 2018
devtools::load_all(pkg = "~/Desktop/datelife")
length(opentree_chronograms$trees) names(opentree_chronograms)
-->
<!--How many trees? How many dated trees?-->
<!-- Methods to include living and fossil lineages are in continued development and increased -->
<!-- usage by the community<!--total evidence and revbayes (fossilized birth-death)-->, -->
<!-- which coupled to better sharing data practices, are greatly contributing to the -->
<!-- accumulation in number and type of available data on taxon ages. -->

>>>>

Phylomatic [@webb2005phylomatic]
<!-- In a way, phylomatic synthesises trees based on a query. -->
is a platform that synthesizes a tree containing taxa of interest from a user supplied set of trees or from its own database.
It has the advantage that it allows grafting taxa of interest absent from their database, but it is limited to mostly seed plants and mammals. If your query does not match its records, it gives many polytomies back.

treePL, is a semi-parametric, rate-smoothing, penalized likelihood dating method
[@Smith2012].
<!--INTRODUCTION: 1st paragraph: general statement that the field knows to be true and important-->
Time of lineage divergence <!--alt lineages: named biological entities, taxon ages?-->
constitutes a fundamental piece of information for evolutionary
understanding in many areas of research, from developmental to conservation biology [@Felsenstein1985a; @Webb2000]. This information can be obtained from fossils and from molecular dating studies. The developments in DNA sequencing and phylogenetic dating inferences have encouraged/spurred the application of molecular dating methods to a very large diversity of organisms, greatly increasing the quantity 
<!-- and improving the quality  (do we know this? we would expect thatm but there is no analysis on that really...)-->
of data on taxon ages across the tree of life.
Along with phylogenetic relationships, it allows testing alternative
evolutionary hypothesis in many areas of research, from developmental to conservation biology [@Felsenstein1985a; @Webb2000].
It has been shown that only a bit of information on time of divergence can improve inferences of phylogenetic distance @Webb2008 bladj
<!--comparative studies, in general, for which knowing
the time context for all life is crucial.-->
Together with geographical distribution patterns, it is a crucial component of historical biogeography studies [@posadas2006historical].
Coupled to species number, it allows studying the tempo and mode of speciation and extinction,
<!-- crucial/vital/critical/essential -->
central for understanding how biodiversity
patterns are shaped across space, time, and clades [@Morlon2014].
<!-- Efforts to have a whole tree of life have been great and here are some examples. -->
<!--INTRODUCTION: 2nd paragraph: important advances in the field-->
In the past two decades, the possibility to obtain good quality DNA sequences
<!-- from a wide variety of organisms became a reality, which,  -->
coupled to methodological developments in phylogenetic and dating inference, allowed
the application of molecular dating methods to a very large diversity
of organisms, greatly increasing the quantity and quality of<!--organisms with--> data on taxon ages across the tree of life.
<!-- figure showing growth on studies with age data from references table from ttol-->
To date, there is a large amount of
<!-- both fossil and molecular-based  -->
data on taxon ages and phylogenetic relationships in public repositories such as Dryad, TreeBASE
and Open Tree of Life (OToL).
<!-- pbdb ??-->
<!--OToL alone hosts now more than 200 chronograms. Cache updated jan 15 2018
devtools::load_all(pkg = "~/Desktop/datelife")
length(opentree_chronograms$trees) names(opentree_chronograms)
-->
<!--How many trees? How many dated trees?-->
<!-- Methods to include living and fossil lineages are in continued development and increased -->
<!-- usage by the community<!--total evidence and revbayes (fossilized birth-death)-->, -->
<!-- which coupled to better sharing data practices, are greatly contributing to the -->
<!-- accumulation in number and type of available data on taxon ages. -->
